Generality in food webs scales with species richness, not latitude

### 2015-01-27 manuscript/latitudepaper.tex
updated results

%Food webs are a thing. (Mention gen, vul, LS.) %Food webs have strong scaling structure.
Food webs --networks of feeding links between species-- have been used for over a century to summarize the structure of 
ecological communities \citep{Williams2000,earlierwork} and to understand how that structure relates to environmental variables 
such as land use \citep{Townsend1998,Townsend2004c,Digel2014}, primary productivity \citep{}, and temperature. The latter
variables in turn have strong gradients over latitude, with productivity and temperature both being higher in the tropics \citep{}, 
suggesting that latitude may indirectly affect food-web structure.
Indeed, these latitudinal environmental gradients have been put forward as potential driverd for the latitudinal gradient in species 
richness, one of the most general and robust patterns in ecology \citep{Schemske2009a,,}.

This effect of latitude on species richness, however, presents a problem for the analysis of other aspects of food-web structure.
Although it was initially thought that food-web structure was small-world and scale invariant \citep{}, further 
analyses with higher-quality data have revealed that most food-web structural properties scale strongly with species 
richness, connectance (the proportion of possible links that are realized), or both \citep{Riede2010} 
(see \citet{Dunne2006} for a review of the history of scaling laws in food web theory). This means that any relationship
between latitude and food-web structure must be considered in light of the latitudinal gradient in species richness.
This is especially true for food-web properties that both scale with species richness and are related to candidate
explanations for the latitudinal diversity gradient, such as measures of specialization [[DBS requests citation, ARC thinks this is the motivation for the paper and isn't aware of a source]].

@@

We then assessed
the impact of latitude on this relationship; specifically, the effect of latitude on the scaling exponent $\beta$. 
In addition to latitude, we included a factorial variable for ecosystem type (stream, lake, terrestrial, marine, or estuary) to account for recognized variation in food web structure across ecosystem types \citep{}.
% We also considered other trends in the data that could affect relationships between latitude and species richness.
% To account for changing standards in data collection over time \citep{Dunne2006} we included an effect of year of
% publication. As 11 of our XX webs included humans, and therefore might have different properties than other networks,
% we included a factorial effect of the presence of humans. 
Finally, some of our webs describe the same area in different
seasons or different years. To account for likely correlation among these webs, we included a random effect of site ID.

@@

\begin{equation}
\label{PowerLat}
Z_{i}=\alpha S_{i}^{\beta_{0}+\beta_{1}L_{i}+\beta_{3}E_{i}} + P_{i} + \epsilon_{i} 
\end{equation}

where $L_{i}$ is the absolute latitude of food web $i$ (degrees from the equator, regardless of direction),
$E_{i}$ indicates the ecosystem type of food web $i$, and $P_{i}$ is a random effect of 
site ID.

@@

\begin{equation}
\label{LogLat}
log(Z_{i})= log(\alpha)+\beta_{0}log(S_{i}) + \beta_{1}L*log(S_{i}) +\beta_{3}E*log(S_{i})+ P_{i} +\epsilon_{i}
\end{equation}

We fit equation \ref{LogLat}, as well as similar equations for $G$ and $V$ using R \citep{R} function lmer from package lme4
\citep{lmerTest}.
We calculated the AIC of the chosen maximal model, as well as the AIC's of a suite of candidate simplified models.
Simplified models were obtained by systematically removing all possible combinations of terms from the full model.
The term for species richness was not removed from a model if any of the terms affecting the exponent $\beta$ were 
included in the model. The model with the lowest AIC was selected as the best-fitting model. 
This simplification was performed using the
R \citep{R} function dredge from package MuMIn \citep{MuMIn}. We then used the R \citep{R} function lmer
from the package lmerTest \citep{lmerTest} to estimate the standardized effects ($\beta$s ) for each fixed effect in the 
best-fitting models, as well as their corresponding $p$-values.

@@

$S$ and $G$ both increased with increasing latitude, contrary to expected latitudinal gradients (Table \ref{Latlms}). 
$G$ and $V$ were not significantly correlated with latitude.

@@

% For models of $Z$, $G$, and $V$, the assumption of multiplicative log-normal error was better supported (Table \ref{Errorfits}). We therefore use only models with logarithmic form for the remainder of the paper.

% \begin{table}[!t]

% \caption{AIC's for model \ref{Power}, assuming normal error, and model \ref{Loglog}, assuming lognormal error.}
% \label{Errorfits}
% \begin{tabular}{l | l l | c}
% Property & Normal error & Lognormal error & Model selected \\
% \hline
% $Z$ & 1151 &  767 & \ref{Loglog} \\
% $G$ & 1266 & 1037 & \ref{Loglog} \\
% $V$ & 1151 & 767 & \ref{Loglog} \\
% \end{tabular}
% \end{table}

In this data set, the best fit models for species richness and links per species did not include an 
effect of latitude on either the scaling exponent or residual (Table \ref{Bestfits}). 
$G$ increased slightly in more recent webs ($\rho_{2}$=0.005, $p$\textless0.001), but the best-fit models for $V$ and $Z$ did not include a term for year of publication. 
In contrast, $Z$ and $V$ were higher in lake food webs ($\beta_{4}$=0.10, $p$\textless0.001 and $\beta_{4}$=0.10, $p$\textless0.001, respectively) while the best-fit model for $G$ did not include an effect for any ecosystem type.

@@

\caption{Standard effects and their $p$-values for the best-fitting models of $Z$, $G$, and $V$.}
\label{Bestfits}
\begin{tabular}{l | l l | l l | l l | l l }
\multirow{2}{*}{Property} & \multicolumn{2}{c}{Intercept} & \multicolumn{2}{c}{log(Species)} & \multicolumn{2}{c}{log(Species)$\Lambda$} & \multicolumn{2}{c}{$\Upsilon$} \\
& $\beta$ & $p$-value &  $\beta$ & $p$-value &  $\beta$ & $p$-value &  $\beta$ & $p$-value  \\
\hline
$Z$ & -0.516 & \textless0.001 & 0.662 & \textless0.001 & 0.099 & \textless0.001 & \multicolumn{2}{c}{NA} \\
$G$ & -9.999 & \textless0.001 & 0.547 & \textless0.001 & \multicolumn{2}{c|}{NA} & 0.005 & \textless0.001 \\
$V$ & -0.516 & \textless0.001 & 0.662 & \textless0.001 & 0.099 & \textless0.001 & \multicolumn{2}{c}{NA} \\
\end{tabular}
\end{table}

@@

The tendency of food-web structure to exhibit scaling relationships with both
species richness and connectance has been  well-established \citep{}. As
species richness in particular is also known to vary systematically over
latitude \citep{}, intuitively one might suspect that any relationship
between food-web properties such as generality might be due to the latitudinal
gradient in species richness. Here, we found that while there were latitudinal
gradients in species richness and generality, latitude was not related to
either the scaling relationship between generality and species richness or the
residuals of this scaling relationship.

The latitudinal gradient in species richness in this data set ran in the opposite 
direction to the general trends for (list types of taxa). This may be due to the 
fact that latitudinal gradient in species richness are usually studied within a
taxonomic or trophic group (e.g., birds, vascular plants), whereas food webs 
measure many taxa concurrently. It is possible that while many taxa generally increase
towards the tropics, the number of species across all taxa that co-occur within an 
study site decreases. Alternatively, it is possible that by using trophic-species versions
of our food webs, we have artificially reversed the species-richness gradient. [[re-run analyses
with species-species webs]]. This would imply that, while there may be more species in the tropics,
many of those species tend to have identical predator and prey sets to co-occuring species. In either
case, it does not seem likely that the increase in
generality with latitude that we observed is due to an underlying tendency for
species to pursue a wider variety of prey at higher latitudes. Rather, the
trend in generality can be more parsimoniously explained by the trend in
species richness.

Our findings here add further doubt to the niche-breadth hypothesis, as
narrower niches in the tropics were associated with lower species richness in
this data set. It appears instead that, as more species are available, each
species tends to consume more prey, be consumed by more predators, and have
more links in general (check with figs). This makes intuitive sense as a more
species-rich community is likely to have more species with similar behaviour
and morphology that thereby have overlapping predator or prey sets.

Parallel increases in generality and species richness across latitude have
also been noted in studies of plant-pollinator networks (although in that case
species richness was highest in the tropics) \citep{Schleuning2012}. Although
the authors of that study do not explicitly examine the relationship between
specialization and network size, they do cite increasing plant species
richness as a potential driver for increasing generalization
\citep{Schleuning2012}. Although the scaling of mutualist network structure
does not appear to have been explored as systematically as that of unipartite
food webs, some poperties such as nestedness \citep{Vazquez2004} and XX
\citep{} have been shown to scale with species richness. As such scaling
appears to be common to both antagonistic and mutualistic food webs, and
trends in species richness can explain counterintuitive trends in network
structure \citep{Schleuning2012}, it seems likely that many latitudinal trends
in network structure are at least partly due to trends in species richness.

%Nestedness increases with A+P, increases number of extreme specialists AND number of generalists. Not clear how degree should scale with S according to scaling with nestedness.

\section*{Conclusion}

It was not our initial intention to question the validity of the wide range of
studies exploring variation in network structure across changing environments \citep{}.
However, if those studies have not taken into account changes in species
richness over environmental gradients, it is possible that observed changes in
food-web structure are due only indirectly to the environmental variables of
interest. Effects of environmental variables on network structure should
therefore be interpreted with caution unless effects of changing species
richness are made explicit.

>>>>

%Food webs are a thing. (Mention gen, vul, LS.) %Food webs have strong scaling structure.
Food webs --networks of feeding links between species-- have been used for over a century to summarize the structure of 
ecological communities \citep{Williams2000,earlierwork} and to understand how that structure relates to environmental variables 
such as land use \citep{Townsend1998,Townsend2004c,Digel2014}, primary productivity \citep{}, and climate\citep{}. The latter
variables in turn have strong gradients over latitude, with productivity and temperature both being higher in the tropics while 
climate is more variable at high latitudes \citep{}, 
suggesting that latitude may indirectly affect food-web structure.
Indeed, these latitudinal environmental gradients have been put forward as potential driverd for the latitudinal gradient in species 
richness, one of the most general and robust patterns in ecology \citep{Schemske2009a,,}.

This effect of latitude on species richness, however, presents a problem for the analysis of other aspects of food-web structure.
Although it was initially thought that food-web structure was small-world and scale invariant \citep{}, further 
analyses with higher-quality data have revealed that most food-web structural properties scale strongly with species 
richness, connectance (the proportion of possible links that are realized), or both \citep{Riede2010} 
(see \citet{Dunne2006} for a review of the history of scaling laws in food web theory). This means that any relationship
between latitude and food-web structure must be considered in light of the latitudinal gradient in species richness.
This is especially true for food-web properties such as measures of specialization
that both scale with species richness \citep{Riede2010} and are related to candidate
explanations for the latitudinal diversity gradient \citep{Vazquez2004}.

@@

We then assessed
the impact of latitude on this relationship; specifically, the effect of latitude on the scaling exponent $\beta$. 
In addition to latitude, we included a factorial variable for ecosystem type (stream, lake, terrestrial, marine, or estuary) to account for recognized variation in food web structure across ecosystem types \citep{}.
% We also considered other trends in the data that could affect relationships between latitude and species richness.
% To account for changing standards in data collection over time \citep{Dunne2006} we included an effect of year of
% publication. As 11 of our XX webs included humans, and therefore might have different properties than other networks,
% we included a factorial effect of the presence of humans. 
Note that some of our webs describe the same area in different
seasons or different years. [[removed the site-level random effect because I don't want it to steal latitude information.]]

@@

\begin{equation}
\label{PowerLat}
Z_{i}=\alpha S_{i}^{\beta_{0}+\beta_{1}L_{i}+\beta_{3}E_{i}} + \epsilon_{i} 
\end{equation}

where $L_{i}$ is the absolute latitude of food web $i$ (degrees from the equator, regardless of direction),
$E_{i}$ indicates the ecosystem type of food web $i$.

@@

\begin{equation}
\label{LogLat}
1log(Z_{i})= log(\alpha)+\beta_{0}log(S_{i}) + \beta_{1}L*log(S_{i}) +\beta_{3}E*log(S_{i}) +\epsilon_{i}
\end{equation}

We fit equation \ref{LogLat}, as well as similar equations for $G$ and $V$ using R \citep{R} function lm from package stats
\citep{stats}.
We calculated the AIC of the chosen maximal model, as well as the AIC's of a suite of candidate simplified models.
Simplified models were obtained by systematically removing all possible combinations of terms from the full model.
The term for species richness was not removed from a model if any of the terms affecting the exponent $\beta$ were 
included in the model. The best-fitting model was then determined to be the model with the fewest terms where $\delta$AIC<2.

The model with the lowest AIC was selected as the best-fitting model. 
This simplification was performed using the
R \citep{R} function dredge from package MuMIn \citep{MuMIn}. We then used the R \citep{R} function lm
to estimate the standardized effects ($\beta$s ) for each fixed effect in the 
best-fitting models, as well as their corresponding $p$-values.

@@

\subsection*{Scaling Relationships with S and Latitude}

For relationships of $Z$, $G$, and $V$ to S, the assumption of multiplicative log-normal error was better supported
(Table \ref{Errorfits}). We therefore use only models with logarithmic form for the remainder of the paper.

Contrary to expected latitudinal gradients, $S$ and $G$ both increased with increasing latitude(Table \ref{Latlms}). 
$G$ and $V$ were not significantly correlated with latitude.

\begin{table}[!t]

\caption{AIC's for model \ref{Power}, assuming normal error, and model \ref{Loglog}, assuming lognormal error.}
\label{Errorfits}
\begin{tabular}{l | l l | c}
Property & Normal error & Lognormal error & Model selected \\
\hline
$Z$ & 1151 &  767 & \ref{Loglog} \\
$G$ & 1266 & 1037 & \ref{Loglog} \\
$V$ & 1151 & 767 & \ref{Loglog} \\
\end{tabular}
\end{table}

@@

\subsection*{Effect of Latitude on Scaling with S}

The effect of latitude on scaling relationships varied across ecosystem types (Table~\ref{Bestfits}). 
In lakes and streams, the scaling exponent for species richness increased with latitude (although this 
increase was not significant for lakes and generality). For other types of ecosystems, the scaling 
exponent for species richness did not change significantly with latitude. 
[[Some more description after figures.]]

@@

\caption{Standard effects and their $p$-values for the best-fitting models of links per species ($Z$), generality ($G$), and vulnerability ($V$).}
\label{Bestfits}
\begin{tabular}{l | l l | l l | l l }
\multirow{2}{*}{Effect} & \multicolumn{2}{c}{$Z$} & \multicolumn{2}{c}{$G$} & \multicolumn{2}{c}{$V$}\\
                & $\beta$ & $p$-value &  $\beta$ & $p$-value &  $\beta$ & $p$-value & \\
\hline
Intercept               & -0.561 & \textless0.001 & -0.299 & \textless0.001 & -0.561 & \textless0.001 \\
log(species)            &  0.733 & \textless0.001 &  0.622 & \textless0.001 &  0.733 & \textless0.001 \\
log(species):Lake       & -0.118 & 0.047          & -0.115 & 0.078          & -0.118 & 0.047 \\
log(species):Marine     &  0.051 & 0.040          &  0.120 & \textless0.001 &  0.051 & 0.040 \\
log(species):Stream     & -0.161 & 0.014          & -0.057 & 0.421          & -0.161 & 0.014 \\
log(species):Latitude   & -0.001 & 0.081          & -0.001 & 0.273          & -0.001 & 0.081 \\
log(species):Lat:Lake   &  0.006 & \textless0.001 &  0.006 & \textless0.001 &  0.006 & \textless0.001 \\
log(species):Lat:Stream &  0.004 & 0.018          &  0.003 & 0.059          &  0.004 & 0.018 \\
\hline
\end{tabular}
\end{table}

@@

[[This'll need to be redone]]
% The tendency of food-web structure to exhibit scaling relationships with both
% species richness and connectance has been well-established \citep{}. As
% species richness in particular is also known to vary systematically over
% latitude \citep{}, intuitively one might suspect that any relationship
% between food-web properties such as generality might be due to the latitudinal
% gradient in species richness. Here, we found that while there were latitudinal
% gradients in species richness and generality, latitude was not related to
% either the scaling relationship between generality and species richness or the
% residuals of this scaling relationship.

% The latitudinal gradient in species richness in this data set ran in the opposite 
% direction to the general trends for (list types of taxa). This may be due to the 
% fact that latitudinal gradient in species richness are usually studied within a
% taxonomic or trophic group (e.g., birds, vascular plants), whereas food webs 
% measure many taxa concurrently. It is possible that while many taxa generally increase
% towards the tropics, the number of species across all taxa that co-occur within an 
% study site decreases. Alternatively, it is possible that by using trophic-species versions
% of our food webs, we have artificially reversed the species-richness gradient. [[re-run analyses
% with species-species webs]]. This would imply that, while there may be more species in the tropics,
% many of those species tend to have identical predator and prey sets to co-occuring species. In either
% case, it does not seem likely that the increase in
% generality with latitude that we observed is due to an underlying tendency for
% species to pursue a wider variety of prey at higher latitudes. Rather, the
% trend in generality can be more parsimoniously explained by the trend in
% species richness.

% Our findings here add further doubt to the niche-breadth hypothesis, as
% narrower niches in the tropics were associated with lower species richness in
% this data set. It appears instead that, as more species are available, each
% species tends to consume more prey, be consumed by more predators, and have
% more links in general (check with figs). This makes intuitive sense as a more
% species-rich community is likely to have more species with similar behaviour
% and morphology that thereby have overlapping predator or prey sets.

% Parallel increases in generality and species richness across latitude have
% also been noted in studies of plant-pollinator networks (although in that case
% species richness was highest in the tropics) \citep{Schleuning2012}. Although
% the authors of that study do not explicitly examine the relationship between
% specialization and network size, they do cite increasing plant species
% richness as a potential driver for increasing generalization
% \citep{Schleuning2012}. Although the scaling of mutualist network structure
% does not appear to have been explored as systematically as that of unipartite
% food webs, some poperties such as nestedness \citep{Vazquez2004} and XX
% \citep{} have been shown to scale with species richness. As such scaling
% appears to be common to both antagonistic and mutualistic food webs, and
% trends in species richness can explain counterintuitive trends in network
% structure \citep{Schleuning2012}, it seems likely that many latitudinal trends
% in network structure are at least partly due to trends in species richness.

% %Nestedness increases with A+P, increases number of extreme specialists AND number of generalists. Not clear how degree should scale with S according to scaling with nestedness.

% \section*{Conclusion}

% It was not our initial intention to question the validity of the wide range of
% studies exploring variation in network structure across changing environments \citep{}.
% However, if those studies have not taken into account changes in species
% richness over environmental gradients, it is possible that observed changes in
% food-web structure are due only indirectly to the environmental variables of
% interest. Effects of environmental variables on network structure should
% therefore be interpreted with caution unless effects of changing species
% richness are made explicit.
